\section{System Design}

\subsection{Network Model}

We use PyWiSim's encounter-based extension to model an intermittently connected wireless network. All nodes are normally placed far apart, so no radio neighbors exist by default. An \texttt{EncounterManager} samples pairwise meetings from a Poisson process and temporarily co-locates the selected node pair for a short encounter interval. During that interval, the nodes communicate using the same \texttt{unicast()} primitive provided by the base simulator. This model intentionally abstracts away detailed mobility trajectories and instead focuses on the key DTN property: communication occurs only when two nodes happen to meet. For the packet-loss robustness study, we additionally enable PyWiSim's built-in probabilistic loss model so that transmissions during an encounter may fail randomly.

This abstraction is appropriate for a first implementation of Binary Spray and Wait because the protocol fundamentally reasons about encounters rather than stable end-to-end routes. It also aligns well with the educational design of PyWiSim, which exposes node behavior through event handlers and message passing without forcing a complex simulator stack.

\subsection{Protocol Logic}

Each message is represented as a bundle with four core fields: message identifier, source, destination, and creation time. Each node additionally stores a local integer copy count for every bundle it currently carries. The source injects a new bundle with an initial copy budget $K$.

Our implementation follows the binary variant of Spray and Wait~\cite{spyropoulos2005spraywait}. When two nodes encounter, each first sends a compact summary containing the set of bundle identifiers it currently holds. After receiving the peer's summary, a node processes each bundle it has that the peer lacks:
\begin{itemize}
    \item If the peer is the destination, the node sends a \texttt{DELIVER} message and the bundle is considered delivered.
    \item Otherwise, if the node holds more than one copy, it enters the spray phase and transfers $\lfloor L/2 \rfloor$ copies to the peer, keeping the remaining copies locally.
    \item If the node holds exactly one copy and the peer is not the destination, it waits and does not forward further.
\end{itemize}

The protocol can therefore be understood as operating in two conceptual phases. In the \emph{spray phase}, the protocol tries to distribute a bounded number of independent forwarding opportunities across the network. A node with many copies should not keep all of them, because that would concentrate delivery chances in a single carrier. Instead, it hands part of its copy budget to newly encountered relays so that the bundle can spread spatially through future contacts. In the \emph{wait phase}, once a node has only one remaining copy, it stops replicating and acts conservatively: it carries that final copy until it directly encounters the destination. This structure is the key idea behind Spray and Wait. It preserves the advantage of multi-copy routing while avoiding the unbounded duplication of epidemic forwarding.

Binary splitting is especially attractive because it balances copies quickly without requiring any global state. Suppose a source starts with $K{=}8$ copies. After one useful encounter, the source and relay each hold 4 copies. A later encounter can split one of those 4-copy holders into two 2-copy holders, and so on, until several relays each hold a single delivery opportunity. This creates multiple parallel chances for successful delivery, but the protocol never exceeds the original copy budget. By contrast, if $K{=}1$, the protocol collapses to direct transmission: no spraying is possible, and the source must wait until it meets the destination itself. If $K{=}2$, there can be at most one spray event before the protocol enters the wait phase. These edge cases are useful for interpreting the evaluation because they explain why increasing $K$ changes the delivery process qualitatively rather than just quantitatively.

Another useful way to view the protocol is as a compromise between two extremes. At one extreme, single-copy forwarding is cheap in terms of data overhead but often suffers very poor delay and delivery probability in sparse networks. At the other extreme, epidemic-style dissemination can achieve excellent reachability, but it pays a large replication cost because every encounter can trigger additional copies. Binary Spray and Wait explicitly chooses a middle point: it allocates only a fixed budget of copies, but uses those copies opportunistically to explore multiple future contact paths. This design is particularly well suited to intermittently connected settings where no node can predict a complete end-to-end path in advance.

This design has two useful properties. First, the total number of bundle copies never exceeds the initial budget $K$. Second, nodes that already reached the wait phase forward only to the final destination, which prevents the explosive replication behavior associated with flooding-based DTN schemes.

\subsection{Implementation Details}

The protocol is implemented in \texttt{spray\_wait.py}. Its main node class,
\texttt{BinarySpray\-Wait\-Node}, stores the current bundles and reacts to four
message types: \texttt{ENCOUNTER}, \texttt{SUMMARY}, \texttt{SPRAY}, and
\texttt{DELIVER}. The encounter callback triggers a summary exchange, while the
summary handler decides whether to spray copies or deliver directly to the
destination.

The message flow during an encounter is intentionally simple. First, the encounter event tells both nodes that a communication opportunity exists. Next, each node sends a summary describing which bundle identifiers it already holds. This step prevents redundant spraying to a node that already carries the same bundle. After comparing summaries, a node either delivers the bundle directly, sprays part of its remaining copy budget, or does nothing if no useful action is possible. Because PyWiSim already models encounters as explicit pairwise events, this protocol logic can be expressed cleanly without a separate neighbor-discovery subsystem or routing-table maintenance procedure.

To support a more honest evaluation, the implementation distinguishes between data and control transmissions:
\begin{itemize}
    \item \textbf{Data transmissions} are bundle-carrying actions: \texttt{SPRAY} and \texttt{DELIVER}.
    \item \textbf{Control transmissions} are summary exchanges used to discover whether the peer already has a given bundle.
\end{itemize}

This distinction matters. In early experiments, counting all transmissions together produced a misleading result in which larger copy budgets sometimes appeared to reduce overhead. The reason was that aggressive replication often delivered the bundle sooner, which shortened the run and reduced subsequent control traffic. Separating data and control overhead makes the tradeoff interpretable: larger $K$ should generally increase forwarding cost, even if it also shortens time-to-delivery.

\subsection{Design Choices and Scope}

We intentionally study a single-bundle, lossless-contact setting. This keeps the evaluation focused on the forwarding policy itself rather than on queue management, wireless contention, buffer overflow, or packet loss. The resulting model is not a full DTN stack, but it is sufficient to test the central algorithmic question of this report: how the Binary Spray and Wait copy budget shapes delivery performance and transmission cost as the network grows.
